---
output: github_document
---

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/followup)](https://cran.r-project.org/package=followup)

# Welcome to the *followup* package!

Contact tracing is an essential part of the COVID-19 response in many countries and of infectious diseases in general. Regular (phone) followup of known contacts to ask them if they have developed symptoms improves can improve the performance of the procedure since contacts who develop symptoms can immediately be directed towards triage and testing facilities, isolation and medical care, decreasing the risk for them to infect others. With high numbers of reported cases the number of contacts in many countries becomes difficult to manage. This package provides a way to objectively prioritize contacts to be followed-up according to the probability that they have yet developed symptoms.

__THIS PACKAGE IS UNDER ACTIVE DEVELOPMENT, PLEASE TAKE CONTACT WITH ITS AUTHORS BEFORE ANY USE__

## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("followup")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval = FALSE}
devtools::install_github("ffinger/followup")
```

Note that this requires the package *devtools* installed.

__Please see the vignette for further details__


# What does it do?

The main features of the package include:

- Compute followup priorities for a list of contacts based on potential dates of exposure, incubation period distributions and dates of last follow-up.
- Outputs the probability that the symptoms onset of a contact was between the last follow up and today.


# Resources

## Vignettes

An overview and examples of *followup* are provided in the vignettes:

followup

Contributions are welcome via **pull requests**.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

