---
title: "followup: incubation functions"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
     df_print: tibble
vignette: >
  %\VignetteIndexEntry{Incubation functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7,
  fig.height=5,
  fig.path="figs-overview/")
```



Load environment:

```{r, echo = TRUE, message=FALSE}
library(dplyr)
library(epitrix)
library(distcrete)
library(followup)
```

Set up the incubation period distribution:

```{r, echo = TRUE}
incubation_days <- 0:3
incubation_frequency <- c(1,3,2,1)
incubation_frequency <- incubation_frequency/sum(incubation_frequency)

plot(incubation_days, incubation_frequency, type = "h")
```

# The integrate_incubation function

Make a list of example contacts each having just one exposure date in `dates_exposure`.

```{r, echo = TRUE}
contact_list <- data.frame(
  id_contact = c("a", "b", "c", "d"),
  type_exposure = c("hospital", "funeral", "family", "restaurant"),
  dates_exposure = as.Date(c("2021-01-15", "2021-01-14", "2021-01-10", "2021-01-16"))
)

contact_list
```


```{r}
date_analysis <- as.Date("2021-01-16")

res <- purrr::map(contact_list$dates_exposure, function(x) integrate_incubation(incubation_frequency, date_low = x, date_high = date_analysis))
res <- unlist(res)

res
```

We can compute the results by hand and compare:

```{r}
expected <- c(4/7, 6/7, 1, 1/7)
testthat::expect_equal(res, expected)
```

Should get exactly the same result using the function including a followup when
 1) last follow-up is NA
 2) last follow-up is in before exposure

```{r}
testthat::expect_equal(
  expected,
  integrate_incubation_followup(incubation_frequency, date_exposure = contact_list$dates_exposure, date_analysis = date_analysis, date_last_followup = NA)
  )

testthat::expect_equal(
  expected,
  integrate_incubation_followup(incubation_frequency, date_exposure = contact_list$dates_exposure, date_analysis = date_analysis, date_last_followup = as.Date("2021-01-01"))
  )
```


# Check the integrate_incubation_followup function

Now add a date of last follow-up for each contact:

```{r}
contact_list$dates_last_followup <- as.Date(c("2021-01-15", "2021-01-15", "2021-01-12", "2021-01-16"))

contact_list
```

```{r}
res <- integrate_incubation_followup(incubation_frequency, date_exposure = contact_list$dates_exposure, date_analysis = date_analysis, date_last_followup = contact_list$dates_last_followup)
```

```{r}
expected <- c((3/7) / (1-1/7), (2/7) / (1-1/7-3/7), 1, 0)
testthat::expect_equal(res, expected)
```

__there is a few more special cases here that should be illustrated here, see code__

